---
title: "Genome-wide ridge regression"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
```

# About

Here I ran genome-wide ridge regression using code implemented at `../methods/gw_ridge`.
In brief, the data is split into 5 folds and we predict 1 fold using ridge regression model trained using the rest 4 folds where the hyperparameter $\theta_g$ is selected via an inner round of cross-validation among the 4 folds of training data.

Here I present the result along with the `OmicKriging` runs done by Owen.

# Load data

My results.

```{r}
df_new = read.table('~/Desktop/tmp/ukb_idp/gw_ridge/gw_ridge_default_theta_g_fold_5_5.tsv.gz', header = T)
```

Owen's runs.

```{r}
nbatch = 444
df_owen = list()
for(i in 1 : nbatch) {
  df_owen[[length(df_owen) + 1]] = read.table(paste0('~/Desktop/tmp/ukb_idp/owen_okriging/okriging_cv_', i, '_summary.txt'), header = T)
}
df_owen = do.call(rbind, df_owen)
df_owen$gene = stringr::str_replace(df_owen$gene, '\\.', '-')
```

Merge the two.

```{r}
df_merge = inner_join(df_new, df_owen, by = c('phenotype' = 'gene'))
```

# Pairwise comparison

```{r, fig.height=5, fig.width=15}
p1 = df_merge %>% ggplot() + geom_point(aes(x = r_squared, y = R2), alpha = 0.2) + 
  th + 
  geom_abline(intercept = 0, slope = 1) +
  xlab('Owen\'s run R2') +
  ylab('New run R2') + 
  coord_equal()

p2 = df_merge %>% ggplot() + geom_point(aes(x = pearson_r, y = Pearson), alpha = 0.2) + 
  th + 
  geom_abline(intercept = 0, slope = 1) +
  xlab('Owen\'s run Pearson Cor') +
  ylab('New run Pearson Cor') + 
  coord_equal()

p3 = df_merge %>% ggplot() + geom_point(aes(x = spearman_r, y = Spearman), alpha = 0.2) + 
  th + 
  geom_abline(intercept = 0, slope = 1) +
  xlab('Owen\'s run Spearman Cor') +
  ylab('New run Spearman Cor') + 
  coord_equal()
p1 + p2 + p3
```


**Take-away**: 

* The new run is consistent with the previous result. Namely, an IDP with good result before also has good result in new run.
* The new run has slighly better performance since we are selecting $\theta_g$ from a grid of values: 0.01, 0.05, 0.1, 0.2, ..., 0.9. But in Owen's run $\theta_g = 1$ which is not ridge regression anymore.


