---
title: "Individual-level ImageXcan (round 1): Fluid intelligence"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://gist.githubusercontent.com/liangyy/4c647634fe00b3f042ebd1599dda65c7/raw/9977562b65d0fb63a693fa7fa60035a37641ad2f/qqplot_by_group')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(data.table)
options(datatable.fread.datatable = F)
source('rlib.R')
library(pander)
panderOptions("table.split.table", Inf)
library(ggpubr)
```

# About

Here we focus on fluid intelligence phenotype in UK Biobank.
Specifically, we extract information from [20016](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=20016) and [20191](https://biobank.ndph.ox.ac.uk/showcase/field.cgi?id=20191).
The phenotype query was done by Sabrina and the resulting files are at `/gpfs/data/im-lab/nas40t2/Data/UKB/ukbrest-queries/2020-11-24-fluid-intelligence`.
The proprocessing of the raw query was done by `misc_data/imagexcan_preprocessing_intelligence.py`.
And the ImageXcan run was done in `submission/imagexcan/run_round_1_fintel.screen`

# Load data

Load dMRI and T1 IDP annotations.

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
dmri = readRDS('../misc_data/download_some_matching_files/annot_dmri_idps.rds')
dmri = dmri %>% mutate(IDP = paste0('IDP-', FieldID))
idps = c('t1', 'dmri')
df_annot = rbind(
  t1 %>% select(IDP, Field),
  dmri %>% select(IDP, Field)
)
```

Load ImageXcan results.

```{r}
df_xcan = list()
for(idp in idps) {
  tmp = read.csv(paste0('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.fintel.linear.', idp, '.csv'))
  tmp2 = read.csv(paste0('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.fintel.susie.', idp, '.csv'))
  tmp = inner_join(tmp, tmp2, by = c('IDP', 'phenotype'))
  df_xcan[[length(df_xcan) + 1]] = tmp %>% mutate(type = idp)
}
df_xcan = do.call(rbind, df_xcan)
```

# Overall QQ-plot

```{r, fig.width=6, fig.height=4}
qqplot_by_group(df_xcan$pval, df_xcan$type) + th
```

# Overall PIP

```{r, fig.width=6, fig.height=4}
df_xcan %>% ggplot() + geom_point(aes(x = IDP, y = pip, color = factor(cs))) + 
  facet_wrap(~type, scales = 'free_x', ncol = 1) + 
  th2 + 
  theme(axis.text.x = element_blank())
```

# Bonferroni significant hits

```{r}
df_xcan %>% mutate(p_adj = pval * n()) %>% filter(p_adj < 0.05) %>% inner_join(df_annot, by = 'IDP') %>% pander(caption = 'Bonferroni significant IDP-level (alpha = 0.05)') 
```


# PIP > 0.25 or within 95% CS

```{r}
df_xcan %>% filter(pip > 0.25) %>% inner_join(df_annot, by = 'IDP') %>% pander(caption = 'PIP > 25%') 
```


```{r}
df_xcan %>% filter(cs > 0) %>% inner_join(df_annot, by = 'IDP')  %>% pander(caption = 'Within 95% credible set') 
```

# Compose the list of IDPs to run MR

There are two openGWAS datasets of interest:

* [ebi-a-GCST006250](https://gwas.mrcieu.ac.uk/datasets/ebi-a-GCST006250/)
* [ukb-b-5238](https://gwas.mrcieu.ac.uk/datasets/ukb-b-5238/)

We focus on the ones within 95% CS along with 25004, 25006, 25008, and 25010 since they appear in a recent [paper](https://www.nature.com/articles/s41467-020-19378-5) analyzing the genetic basis of intelligence and brain volume. 
Note that this is NOT a direct replication of their work since they use the sum of 25004, 25006, 25008 which do not appear in our IDP set. 
Also, they may use intelligence GWASs other than the one we used.

```{r}
opengwass = data.frame(pheno = c('intelligence', 'fluid_intelligence'), pheno_code = c('ebi-a-GCST006250', 'ukb-b-5238'))
selected_idps = rbind(
  df_xcan %>% filter(cs > 0) %>% select(IDP, type), 
  data.frame(IDP = paste0('IDP-', 25000 + c(4, 6, 8, 10)), type = 't1')
) %>% distinct()
for(type_ in unique(selected_idps$type)) {
  sub = selected_idps %>% filter(type == type_) 
  tmp = cbind(opengwass, rep(sub$IDP, each = nrow(opengwass)))
  colnames(tmp)[3] = 'idp'
  write.table(tmp, paste0('../misc_data/imagexcan_round_1.fintel.', type_, '.signif.tsv'), quo = F, row = F, col = T, sep = '\t')
}
```

# MR results

Here we only show results that at least one direction has two p-values < 0.05.

```{r}
df = list()
for(type_ in unique(selected_idps$type)) {
  df[[length(df) + 1]] = read.table(paste0('../misc_data/imagexcan_round_1.fintel.', type_, '.signif.tsv'), header = T) %>% mutate(type = type_)
}
df = do.call(rbind, df)
collector_mr = list()
collector_data = list()
for(i in 1 : nrow(df)) {
  tmp = readRDS(paste0('~/Desktop/tmp/ukb_idp/mr/MR.fintel_', df$type[i], '.', df$idp[i], '_x_', df$pheno[i], '.rds'))
  if(sum(tmp$idp2pheno$mr$pval < 0.05) > 2 | sum(tmp$pheno2idp$mr$pval < 0.05) > 2) {
    collector_mr[[length(collector_mr) + 1]] = rbind(
      tmp$idp2pheno$mr %>% mutate(direction = 'idp2pheno', pheno = df$pheno[i], idp = df$idp[i]),
      tmp$pheno2idp$mr %>% mutate(direction = 'pheno2idp', pheno = df$pheno[i], idp = df$idp[i])
    )
    collector_data[[length(collector_data) + 1]] = rbind(
      tmp$idp2pheno$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'idp2pheno', pheno = df$pheno[i], idp = df$idp[i]),
      tmp$pheno2idp$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'pheno2idp', pheno = df$pheno[i], idp = df$idp[i])
    )
  }
}
df_data = do.call(rbind, collector_data) %>% inner_join(df_annot, by = c('idp' = 'IDP'))
df_mr = do.call(rbind, collector_mr) %>% inner_join(df_annot, by = c('idp' = 'IDP'))
```

The selected MR pairs.

```{r}
df_mr %>% select(pheno, idp) %>% distinct() %>% pander
df_tmp = df_mr %>% select(pheno, idp) %>% distinct()
```

Show the plot.

```{r, fig.width=15, fig.height=5}
for(i in 1 : nrow(df_tmp)) {
  mydat = df_data %>% filter(pheno == df_tmp$pheno[i], idp == df_tmp$idp[i])
  mymr = df_mr %>% filter(pheno == df_tmp$pheno[i], idp == df_tmp$idp[i])
  idp_name = mydat$Field[1]
  gwas_name = mydat$pheno[1]
  plist = list()
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'idp2pheno') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(idp_name, ' -> \n ', gwas_name)) +
    theme(legend.position = 'none')
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'pheno2idp') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(gwas_name, ' -> \n ', idp_name)) +
    theme(legend.position = 'none')
  # plist[[length(plist) + 1]] = gridExtra::tableGrob(df_mr %>% select(method, nsnp, b, pval, direction))
  qq = ggarrange(plist[[1]], plist[[2]], ggtexttable(format(mymr %>% select(method, nsnp, b, pval, direction), digits = 3)), ncol = 3)
  print(qq)
}

```


