---
title: "Individual-level ImageXcan (round 1) T1"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(data.table)
options(datatable.fread.datatable = F)
pheno_interest = c('weekly_alcohol', 'recurrent_depressive_disorder', 'parent_depression', 'parent_AD', 'handedness', 'daily_coffee', 'daily_cigarettes')
pheno_bcc = c('wbc', 'rbc', 'platelet', 'lymphocyte', 'monocyte', 'neutrophil', 'eosinophil', 'basophil')
pheno_ht = c('dbp', 'sbp') # , 'ht', 'hb', 'mcv', 'mch', 'mchc')
pheno_rand = paste0('random_pheno_', 0 : 9)
pheno_sig_dmri = paste0('single_sig_dmri_', 0 : 9)
pheno_sig_orig = paste0('single_sig_orig_', 0 : 9)
source('rlib.R')
```

# About

See running setup at 

* The original IDPs (residual obtained by Owen): `submission/imagexcan/run_round_1_linear.orignal.screen`.
* T1 IDP adjusted for first 5 PCs: `submission/imagexcan/run_round_1_linear.dmri.screen`.

The current results are generated by linear regression. 
The logistic regression is ready to go but it has some numerical issue when including too many covariates.

We have special focus on T1 IDPs.

# Load data

T1 IDPs.

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
```

Load results with original IDPs and adjusted IDPs.

```{r}
# linear solver
df_lm = read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.linear.original.csv') %>% filter(IDP %in% t1$IDP)
df_lm_resid = read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.linear.t1.csv') %>% filter(IDP %in% t1$IDP)
df_lm = rbind(
  df_lm %>% mutate(idp_type = 'original'),
  df_lm_resid %>% mutate(idp_type = 'regress_out')
)

df_lm = df_lm %>% left_join(t1 %>% select(IDP, normalized_by_head_size, position, matter_type, lr), by = 'IDP')
df_lm$lr[ is.na(df_lm$lr) ] = 'NA'
df_lm$matter_type[ is.na(df_lm$matter_type) ] = 'NA'
```


# Overall QQ-plot

```{r, fig.width=16, fig.height=12}
p1 = df_lm %>% filter(phenotype %in% pheno_interest) %>% group_by(phenotype, idp_type) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% 
  ggplot() + 
  geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0)
p2 = df_lm %>% filter(phenotype %in% pheno_bcc) %>% group_by(phenotype, idp_type) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% 
  ggplot() + 
  geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0)
p3 = df_lm %>% filter(phenotype %in% pheno_ht) %>% group_by(phenotype, idp_type) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% 
  ggplot() + 
  geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0)
p4 = df_lm %>% filter(phenotype %in% pheno_rand) %>% group_by(phenotype, idp_type) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% 
  ggplot() + 
  geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0)

(p1 + p2) / (p3 + p4) 
```

# Bonferroni significant hits

```{r}
df_lm %>% filter(idp_type == 'regress_out') %>% mutate(p_adj = pval * n()) %>% filter(p_adj < 0.05, phenotype %in% pheno_interest) %>% pander::pander(caption = 'Bonferroni significant IDP-level (alpha = 0.05) among all focal traits (adjusted IDP)')
df_lm %>% filter(idp_type == 'original') %>% mutate(p_adj = pval * n()) %>% filter(p_adj < 0.05, phenotype %in% pheno_interest) %>% pander::pander(caption = 'Bonferroni significant IDP-level (alpha = 0.05) among all focal traits (original IDP)')
```

# Susier run

```{r}
df_susie = read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.susie.t1.csv') # %>% filter(IDP %in% dmri$IDP)
df_susie = df_susie %>% left_join(t1 %>% select(IDP, normalized_by_head_size, position, matter_type, lr), by = 'IDP')
df_susie$lr[ is.na(df_susie$lr) ] = 'NA'
df_susie$matter_type[ is.na(df_susie$matter_type) ] = 'NA'
```

# PIP plot

```{r, fig.width=16, fig.height=12}
p1 = df_susie %>% filter(phenotype %in% pheno_interest) %>% 
  ggplot() + 
  geom_point(aes(x = IDP, y = pip, color = factor(cs))) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0) +
  theme(axis.text.x = element_blank())
p2 = df_susie %>% filter(phenotype %in% pheno_bcc) %>%
  ggplot() + 
  geom_point(aes(x = IDP, y = pip, color = factor(cs))) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0) +
  theme(axis.text.x = element_blank())
p3 = df_susie %>% filter(phenotype %in% pheno_ht) %>%
  ggplot() + 
  geom_point(aes(x = IDP, y = pip, color = factor(cs))) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0) +
  theme(axis.text.x = element_blank())
p4 = df_susie %>% filter(phenotype %in% pheno_rand) %>% 
  ggplot() + 
  geom_point(aes(x = IDP, y = pip, color = factor(cs))) + 
  facet_wrap(~phenotype, scales = 'free_y') + th2 + geom_abline(slope = 1, intercept = 0) +
  theme(axis.text.x = element_blank())

(p1 + p2) / (p3 + p4)
```

# 95% credible set

```{r}
df_susie %>% filter(phenotype %in% pheno_interest) %>% filter(cs > 0) %>% pander::pander(caption = 'IDP in 95% credible set among all focal phenotypes')
```

# PIP > 0.25

```{r}
df_susie %>% filter(phenotype %in% pheno_interest) %>% filter(pip > 0.25) %>% pander::pander(caption = 'IDP in 95% credible set among all focal phenotypes')
```

# Obtaining the list of significant pairs (95% CS or PIP > 0.25)

```{r}
tmp = rbind(
  df_susie %>% filter(phenotype %in% pheno_interest) %>% filter(cs > 0),
  df_susie %>% filter(phenotype %in% pheno_interest) %>% filter(pip > 0.25)
)
df_gwas = read.csv('../misc_data/pheno_of_interest.meta_table.csv')
tmp = tmp %>% select(IDP, phenotype) %>% distinct()
tmp = left_join(tmp, df_gwas, by = 'phenotype')
tmp %>% select(id, IDP, gwas_code) %>% rename(pheno = id, idp = IDP, pheno_code = gwas_code) %>% write.table('../misc_data/imagexcan_round_1.t1.signif.tsv', sep = '\t', quo = F, row = F, col = T)
```

