---
title: "Round 2: BrainXcan results of UKB phenotypes"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://raw.githubusercontent.com/liangyy/misc-tools/master/plot_tool/plot_tools.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(ggrepel)
library(data.table)
library(ggpubr)
options(datatable.fread.datatable = F)
source('rlib.R')
pheno_interest = c('weekly_alcohol', 'recurrent_depressive_disorder', 'parent_depression', 'parent_AD', 'handedness', 'daily_coffee', 'daily_cigarettes')
pheno_bcc = c('wbc', 'rbc', 'platelet', 'lymphocyte', 'monocyte', 'neutrophil', 'eosinophil', 'basophil')
pheno_ht = c('dbp', 'sbp', 'ht', 'hb', 'mcv', 'mch', 'mchc')
pheno = c(pheno_interest, pheno_bcc, pheno_ht)
```

# About

We run individual-level BrainXcan on some UKB phenotypes.
The IDPs being used are listed below:

* `scaled_t1_all_covar_w_pc`
* `orignal_dmri_all_covar_w_pc`

# Load data

```{r}
models = c('ridge', 'en')
idps = c('dmri','t1') #  
# types = c('linear', 'susie')
df = list()
for(m in models) {
  for(t in idps) {
    fn1 = paste0('~/Desktop/tmp/ukb_idp/data/imagexcan_round_2.linear.', t, '_', m, '.csv')
    tmp1 = read.csv(fn1)
    fn2 = paste0('~/Desktop/tmp/ukb_idp/data/imagexcan_round_2.susie.', t, '_', m, '.csv')
    tmp2 = read.csv(fn2)
    tmp = inner_join(tmp1, tmp2, by = c('IDP', 'phenotype'))
    df[[length(df) + 1]] = tmp %>% mutate(model = m, idp_type = t)
  }
}
df = do.call(rbind, df)
df$bhat = - df$bhat
df_all = df
df = df_all %>% filter(phenotype %in% pheno_interest)
```

Load IDP annotations. 

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
dmri = readRDS('../misc_data/download_some_matching_files/annot_dmri_idps.rds')
dmri = dmri %>% mutate(IDP = paste0('IDP-', FieldID))
idps = c('t1', 'dmri')
df_annot = rbind(
  t1 %>% select(IDP, Field) %>% mutate(idp_type = 't1'),
  dmri %>% select(IDP, Field) %>% mutate(idp_type = 'dmri')
)
```

# Overview of the results

```{r, fig.height=4, fig.width=5}
p1 = qqplot_by_group(df$pval, paste(df$model, df$idp_type)) + th + ggtitle('QQ-plot: UKB phenotypes of interest') 
p1
```


```{r, fig.height=4, fig.width=8}
p1 = df %>% reshape2::dcast(IDP + phenotype + idp_type ~ model, value.var = 'pip') %>% ggplot() +
  geom_point(aes(x = ridge, y = en), alpha = 0.05) +
  th2 + ggtitle('QQ-plot: UKB phenotypes of interest') + facet_wrap(~idp_type)
p1 
```



# Selecting putative signals

Meet both criteria:

* Bonferroni significant at $\alpha = 0.01$ (we used stringent cutoff since there is some sort of inflation).
* Within 95% credible set.

```{r}
df_signif = df %>% group_by(idp_type, model) %>% mutate(p_adj = pval * n()) %>% ungroup() %>%
  filter(cs > 0 & p_adj < 0.01)

df_signif %>% group_by(idp_type, model) %>% summarize(nsig = n()) %>% pander::pander(caption = 'The number of putative signals extracted from UKB phenotypes of interest')
```

Save putative signals to TSV.

```{r}
df_gwas = read.csv('../misc_data/pheno_of_interest.meta_table.csv')
for(dd in c('dmri', 't1')) {
  sub = df_signif %>% filter(idp_type == dd) %>% select(IDP, phenotype) %>% distinct()
  sub = left_join(sub, df_gwas, by = 'phenotype')
  sub = sub %>% select(id, IDP, gwas_code) %>% rename(pheno = id, idp = IDP, pheno_code = gwas_code)
  write.table(sub, paste0('../misc_data/round_2_selected.ukb_pheno.', dd, '.signif.tsv'), quo = F, col = T, row = F, sep = '\t')
}
```

# Mendelian randomization results

```{r}
collector_mr = list()
collector_data = list()
idps = c('t1', 'dmri') 
df_signif_out = left_join(df_signif, df_gwas, by = 'phenotype') %>% select(idp_type, IDP, phenotype, id) %>% distinct()
for(dd in idps) {
  sub = df_signif_out %>% filter(idp_type == dd) %>% select(-idp_type)
  for(i in 1 : nrow(sub)) {
    tmp = readRDS(paste0('~/Desktop/tmp/ukb_idp/mr_ukb_pheno_2nd/MR.ukb_pheno_2nd_', dd, '.', sub$IDP[i], '_x_', sub$id[i], '.rds'))
    if(sum(is.na(tmp$idp2pheno$mr)) == 1 | sum(is.na(tmp$pheno2idp$mr)) == 1) {
      next
    }
    n_idp2pheno = sum(tmp$idp2pheno$mr$pval < 0.05) 
    n_pheno2idp = sum(tmp$pheno2idp$mr$pval < 0.05) 
    if(n_idp2pheno >= 2 | n_pheno2idp >= 2) {
      collector_mr[[length(collector_mr) + 1]] = rbind(
        tmp$idp2pheno$mr %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd, id = sub$id[i]),
        tmp$pheno2idp$mr %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd, id = sub$id[i])
      )
      collector_data[[length(collector_data) + 1]] = rbind(
        tmp$idp2pheno$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd, id = sub$id[i]),
        tmp$pheno2idp$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd, id = sub$id[i])
      )
    } 
  }
}

df_data = do.call(rbind, collector_data) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr = do.call(rbind, collector_mr) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr$Field[ is.na(df_mr$Field) ] = paste(df_mr$idp_type[ is.na(df_mr$Field) ], df_mr$IDP[ is.na(df_mr$Field) ])
df_data$Field[ is.na(df_data$Field) ] = paste(df_data$idp_type[ is.na(df_data$Field) ], df_data$IDP[ is.na(df_data$Field) ])
```


```{r, fig.width=15, fig.height=5}
df_all_signif = df
df_all_signif = df_all_signif %>% rename(cs95 = cs)
plot_pair = function(pheno_, idp_, type_) {
  mydat = df_data %>% filter(id == pheno_, IDP == idp_, idp_type == type_)
  mymr = df_mr %>% filter(id == pheno_, IDP == idp_, idp_type == type_)
  idp_name = mydat$Field[1]
  gwas_name = mydat$pheno[1]
  idp_id = mydat$IDP[1]
  this_result = df_all_signif %>% filter(IDP == idp_id, phenotype == gwas_name, idp_type == type_)
  plist = list()
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'idp2pheno') %>% ggplot() + 
    geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + 
    geom_vline(xintercept = 0, color = 'lightgray') + 
    geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(idp_name, ' -> \n ', gwas_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = bhat, intercept = 0, color = model))
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'pheno2idp') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(gwas_name, ' -> \n ', idp_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = 1 / bhat, intercept = 0, color = model))
  # plist[[length(plist) + 1]] = gridExtra::tableGrob(df_mr %>% select(method, nsnp, b, pval, direction))
  this_result2 = this_result %>% mutate(method = paste0('imgx_', model), nsnp = paste0('pip = ', signif(pip, digits = 2), '\n cs = ', cs95), b = bhat, direction = NA) 
  mymr = rbind(mymr %>% select(method, nsnp, b, pval, direction), this_result2 %>% select(method, nsnp, b, pval, direction))
  qq = ggarrange(plist[[1]], plist[[2]], ggtexttable(format(mymr, digits = 3)), ncol = 3)
  qq
}
```


```{r, fig.width=15, fig.height=5}
df_tmp = df_mr %>% select(id, IDP, idp_type) %>% distinct()  # %>% filter(is_pass == T) 
for(i in 1 : nrow(df_tmp)) {
  qq = plot_pair(df_tmp$id[i], df_tmp$IDP[i], df_tmp$idp_type[i])
  print(qq)
}

```


# Compare to S-BrainXcan

```{r, fig.height=4, fig.width=15}
# get the mapping between the two phenotype sets
df_map = data.frame(
  id = c('wbc', 'rbc', 'platelet', 'lymphocyte', 'monocyte', 'neutrophil', 'eosinophil', 'height', 'bmi'),
  gwas = c('Astle_et_al_2016_White_blood_cell_count', 'Astle_et_al_2016_Red_blood_cell_count', 'Astle_et_al_2016_Platelet_count', 'Astle_et_al_2016_Lymphocyte_counts', 'Astle_et_al_2016_Monocyte_count', 'Astle_et_al_2016_Neutrophil_count', 'Astle_et_al_2016_Eosinophil_counts', 'UKB_50_Standing_height', 'UKB_21001_Body_mass_index_BMI')
)

df_gg = list()
models = c(ridge = 'ridge', en = 'elastic_net')
idps = c(t1 = 't1.scaled', dmri = 'dmri.original')
for(i in df_map$gwas) {
  for(m in names(models)) {
    for(ii in names(idps)) {
      df_gg[[length(df_gg) + 1]] = read.csv(paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_gtex_gwas_2nd/', idps[[ii]], '.all_covar.w_pc.gw_', models[[m]], '_beta_x_', i, '_x_simagexcan.csv')) %>% mutate(model = m, idp_type = ii, phenotype = i)
    }
  }
}
df_gg = do.call(rbind, df_gg)
df_gg = df_gg %>% rename(bhat.sbxcan = bhat, pval.sbxcan = pval, pip.sbxcan = pip, cs.sbxcan = cs95)

df2 = df_map %>% inner_join(
  df_all %>% rename(bhat.bxcan = bhat, pval.bxcan = pval, pip.bxcan = pip, cs.bxcan = cs), 
  by = c('id' = 'phenotype')
)
df2 = df2 %>% inner_join(df_gg, by = c('gwas' = 'phenotype', 'IDP', 'idp_type', 'model'))
df2 = df2 %>% mutate(zscore.bxcan = p2z(pval.bxcan, bhat.bxcan), zscore.sbxcan = p2z(pval.sbxcan, bhat.sbxcan))

df2 %>% ggplot() + geom_point(aes(x = zscore.bxcan, y = zscore.sbxcan, color = idp_type)) + facet_grid(model~id) + th2 + coord_equal() + geom_abline(slope = 1, intercept = 0)

df2 %>% ggplot() + geom_point(aes(x = pip.bxcan, y = pip.sbxcan, color = idp_type), alpha = 0.5) + facet_grid(model~id) + th2 + coord_equal() + geom_abline(slope = 1, intercept = 0) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
