---
title: "Round 2: S-BrainXcan vs genetic correlation"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://raw.githubusercontent.com/liangyy/misc-tools/master/plot_tool/plot_tools.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(ggrepel)
library(data.table)
options(datatable.fread.datatable = F)
source('rlib.R')
library(VennDiagram)
library(ggpubr)
```

# About

Here we present the first set of genetic correltaion results between IDPs and phenotypes of interest. We compare the genetic correlation results to the S-BrainXcan.
We considered GTEx-GWAS phenotypes and the psychiatric phenotypes

# Load data

## Load GTEx-GWAS results.

Load S-BrainXcan.

```{r}
traits = read.table('../submission/simagexcan/gtex_gwas_list.txt')$V1
models = c('ridge', 'elastic_net')
tags = list(dmri = 'dmri.original.all_covar.w_pc', t1 = 't1.scaled.all_covar.w_pc')
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    for(mm in models) {
      tmp = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_gtex_gwas_2nd/', tags[[nn]], '.gw_', mm, '_beta_x_', rr, '_x_simagexcan.csv')
      df[[length(df) + 1]] = read.csv(tmp) %>% 
              mutate(idp_type = nn, phenotype = rr, model = mm)
    }
  }
}

df = do.call(rbind, df)
df_sb1 = df %>% mutate(zscore = p2z(pval, bhat))
# df = df %>% filter(is.na(stringr::str_match(IDP, 'PC')[, 1]))
```


Load genetic correlation results.

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/genetic_cor_2nd/', nn, '_2nd_x_gtex-gwas_x_', rr, '.ldsc_rg.log')
    tmp = load_ldsc_rg(tmp)
    tmp = tmp %>% select(p2, rg, p, z, h2_obs, h2_obs_se) %>% rename(IDP = p2, pval = p, zscore = z)
    df[[length(df) + 1]] = tmp %>% 
      mutate(idp_type = nn, phenotype = rr)
  }
}
df_cor1 = do.call(rbind, df)
```

## Load the psychiatric phenotypes

Load S-BrainXcan.

```{r}
traits = read.table('../misc_data/preprocess_psychiatric_traits/trait_list.txt')$V1
models = c('ridge', 'elastic_net')
tags = list(dmri = 'dmri.original.all_covar.w_pc', t1 = 't1.scaled.all_covar.w_pc')
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    for(mm in models) {
      tmp = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_psychiatric_2nd/', tags[[nn]], '.gw_', mm, '_beta_x_', rr, '_x_simagexcan.csv')
      df[[length(df) + 1]] = read.csv(tmp) %>% 
              mutate(idp_type = nn, phenotype = rr, model = mm)
    }
  }
}

df = do.call(rbind, df)
df_sb2 = df %>% mutate(zscore = p2z(pval, bhat))
# df = df %>% filter(is.na(stringr::str_match(IDP, 'PC')[, 1]))
```

Load genetic correlation results.

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/genetic_cor_2nd/', nn, '_2nd_x_psychiatric_x_', rr, '.ldsc_rg.log')
    tmp = load_ldsc_rg(tmp)
    tmp = tmp %>% select(p2, rg, p, z, h2_obs, h2_obs_se) %>% rename(IDP = p2, pval = p, zscore = z)
    df[[length(df) + 1]] = tmp %>% 
      mutate(idp_type = nn, phenotype = rr)
  }
}
df_cor2 = do.call(rbind, df)
```

## Load h2

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/heritability_2nd_round/', tags[[nn]], '.tsv.gz')
    df[[length(df) + 1]] = read.table(tmp, sep = '\t', header = T) %>% mutate(idp_type = nn) %>% rename(IDP = phenotype)
  }
}
df_h2 = do.call(rbind, df)
```

## Combine 

```{r}
df_sb = rbind(
  df_sb1 %>% mutate(data = 'gtex-gwas'), 
  df_sb2 %>% mutate(data = 'psychiatric')
)
df_cor = rbind(
  df_cor1 %>% mutate(data = 'gtex-gwas'), 
  df_cor2 %>% mutate(data = 'psychiatric')
)
```

## QQ-plot

```{r, fig.width=6, fig.height=3}
df_cor %>% group_by(data, idp_type) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% ungroup() %>%
  ggplot() + geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + facet_wrap(~data) + th2 + 
  geom_abline(slope = 1, intercept = 0, color = 'gray') +
  ggtitle('QQ-plot of LDSC RG')
```


```{r, fig.width=6, fig.height=6}
df_sb %>% group_by(data, idp_type, model) %>% mutate(pexp = rank(pval) / (n() + 1)) %>% ungroup() %>%
  mutate(pval = pmax(pval, 1e-30)) %>%
  ggplot() + geom_point(aes(x = -log10(pexp), y = -log10(pval), color = idp_type)) + facet_grid(model~data) + th2 + 
  geom_abline(slope = 1, intercept = 0, color = 'gray') +
  ggtitle('QQ-plot of S-BrainXcan')
```

# S-BrainXcan vs genetic correlation

```{r, fig.width=6, fig.height=6}
tmp = inner_join(df_sb, df_cor, by = c('phenotype', 'idp_type', 'IDP', 'data'), suffix = c('.SB', '.RG'))
tmp %>% ggplot() + geom_point(aes(x = zscore.RG, y = zscore.SB), alpha = 0.2) + facet_grid(model~data) + th2 + geom_hline(yintercept = 0, color = 'gray') + geom_vline(xintercept = 0, color = 'gray')
```

```{r, fig.width=6, fig.height=6}
# tmp = inner_join(df_sb, df_cor, by = c('phenotype', 'idp_type', 'IDP', 'data'), suffix = c('.SB', '.RG'))
tmp %>% filter(pval.SB < 1e-2) %>% ggplot() + geom_point(aes(x = rg, y = bhat), alpha = 0.2) + facet_grid(model~data) + th2 + 
  scale_y_continuous(lim = c(-1, 1)) +
  geom_hline(yintercept = 0, color = 'gray') + geom_vline(xintercept = 0, color = 'gray')
```

# GCTA h2 vs LDSC h2

```{r}
tmp = inner_join(df_h2, df_cor, by = c('idp_type', 'IDP'), suffix = c('.GCTA', '.LDSC'))
tmp = tmp %>% group_by(IDP, idp_type) %>% 
  summarize(
    h2.LDSC = mean(h2_obs), h2_se.LDSC = max(h2_obs_se), 
    h2.GCTA = h2[1], h2_se.GCTA = h2_SE[1]
  ) %>% ungroup()
tmp %>% ggplot() + geom_point(aes(x = h2.GCTA, y = h2.LDSC)) + 
  geom_errorbar(aes(x = h2.GCTA, ymin = h2.LDSC - 1.96 * h2_se.LDSC, ymax = h2.LDSC + 1.96 * h2_se.LDSC), alpha = 0.2) + 
  geom_errorbarh(aes(y = h2.LDSC, xmin = h2.GCTA - 1.96 * h2_se.GCTA, xmax = h2.GCTA + 1.96 * h2_se.GCTA), alpha = 0.2) +
  facet_wrap(~idp_type) + th2 + geom_abline(slope = 1, intercept = 0, color = 'gray')
```

# MR runs (Psychiatric traits) 

## Extract Bonferroni significant pairs

```{r}
df_signif = df_cor %>% filter(phenotype %in% traits) %>% group_by(idp_type) %>% mutate(p_adj = pval * n()) %>% filter(p_adj < 0.05) %>% ungroup()
for(dd in c('dmri', 't1')) {
  sub = df_signif %>% filter(idp_type == dd) %>% select(IDP, phenotype) %>% distinct()
  sub$pheno2 = sub$phenotype
  colnames(sub) = c('idp', 'pheno', 'pheno_code')
  sub = sub[, c('pheno', 'pheno_code', 'idp')]
  write.table(sub, paste0('../misc_data/round_2_selected_gencor.psychiatric.', dd, '.signif.tsv'), quo = F, col = T, row = F, sep = '\t')
}
```

## Mendelian randomization results

Load IDP annotations. 

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
dmri = readRDS('../misc_data/download_some_matching_files/annot_dmri_idps.rds')
dmri = dmri %>% mutate(IDP = paste0('IDP-', FieldID))
idps = c('t1', 'dmri')
df_annot = rbind(
  t1 %>% select(IDP, Field) %>% mutate(idp_type = 't1'),
  dmri %>% select(IDP, Field) %>% mutate(idp_type = 'dmri')
)
```

```{r}
collector_mr = list()
collector_data = list()
idps = c('t1', 'dmri') 
df_signif_out = df_signif %>% select(idp_type, IDP, phenotype) %>% distinct()
for(dd in idps) {
  sub = df_signif_out %>% filter(idp_type == dd) %>% select(-idp_type)
  for(i in 1 : nrow(sub)) {
    tmp = readRDS(paste0('~/Desktop/tmp/ukb_idp/mr_psychiatric_2nd/MR_local.psychiatric_2nd_gencor_', dd, '.', sub$IDP[i], '_x_', sub$phenotype[i], '.rds'))
    if(sum(is.na(tmp$idp2pheno$mr)) == 1 | sum(is.na(tmp$pheno2idp$mr)) == 1) {
      next
    }
    n_idp2pheno = sum(tmp$idp2pheno$mr$pval < 0.05) 
    n_pheno2idp = sum(tmp$pheno2idp$mr$pval < 0.05) 
    if(n_idp2pheno > 2 | n_pheno2idp > 2) {
      collector_mr[[length(collector_mr) + 1]] = rbind(
        tmp$idp2pheno$mr %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd),
        tmp$pheno2idp$mr %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd)
      )
      collector_data[[length(collector_data) + 1]] = rbind(
        tmp$idp2pheno$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd),
        tmp$pheno2idp$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd)
      )
    } 
  }
}

df_data = do.call(rbind, collector_data) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr = do.call(rbind, collector_mr) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr$Field[ is.na(df_mr$Field) ] = paste(df_mr$idp_type[ is.na(df_mr$Field) ], df_mr$IDP[ is.na(df_mr$Field) ])
df_data$Field[ is.na(df_data$Field) ] = paste(df_data$idp_type[ is.na(df_data$Field) ], df_data$IDP[ is.na(df_data$Field) ])
```


```{r, fig.width=15, fig.height=5}
df_all_signif = df_cor
plot_pair = function(pheno_, idp_, type_) {
  mydat = df_data %>% filter(pheno == pheno_, IDP == idp_, idp_type == type_)
  mymr = df_mr %>% filter(pheno == pheno_, IDP == idp_, idp_type == type_)
  idp_name = mydat$Field[1]
  gwas_name = mydat$pheno[1]
  idp_id = mydat$IDP[1]
  this_result = df_all_signif %>% filter(IDP == idp_id, phenotype == gwas_name, idp_type == type_)
  plist = list()
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'idp2pheno') %>% ggplot() + 
    geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + 
    geom_vline(xintercept = 0, color = 'lightgray') + 
    geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(idp_name, ' -> \n ', gwas_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = bhat, intercept = 0, color = model))
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'pheno2idp') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(gwas_name, ' -> \n ', idp_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = 1 / bhat, intercept = 0, color = model))
  # plist[[length(plist) + 1]] = gridExtra::tableGrob(df_mr %>% select(method, nsnp, b, pval, direction))
  this_result2 = this_result %>% mutate(method = 'gen_cor', nsnp = NA, b = rg, direction = NA) 
  mymr = rbind(mymr %>% select(method, nsnp, b, pval, direction), this_result2 %>% select(method, nsnp, b, pval, direction))
  qq = ggarrange(plist[[1]], plist[[2]], ggtexttable(format(mymr, digits = 3)), ncol = 3)
  qq
}
```


```{r, fig.width=15, fig.height=5}
df_tmp = df_mr %>% select(pheno, IDP, idp_type) %>% distinct()  # %>% filter(is_pass == T) 
for(i in 1 : nrow(df_tmp)) {
  qq = plot_pair(df_tmp$pheno[i], df_tmp$IDP[i], df_tmp$idp_type[i])
  print(qq)
}

```

