---
title: "Round 2: S-BrainXcan vs genetic correlation"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://raw.githubusercontent.com/liangyy/misc-tools/master/plot_tool/plot_tools.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(ggrepel)
library(data.table)
options(datatable.fread.datatable = F)
source('rlib.R')
library(VennDiagram)
library(ggpubr)
```

# About

Here we present the first set of genetic correltaion results between IDPs and phenotypes of interest. We compare the genetic correlation results to the S-BrainXcan.
We considered GTEx-GWAS phenotypes and the psychiatric phenotypes

# Load data

## Load GTEx-GWAS results.

Load S-BrainXcan.

```{r}
traits = read.table('../submission/simagexcan/gtex_gwas_list.txt')$V1
models = c('ridge', 'elastic_net')
tags = list(dmri = 'dmri.original.all_covar.w_pc', t1 = 't1.scaled.all_covar.w_pc')
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    for(mm in models) {
      tmp = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_gtex_gwas_2nd/', tags[[nn]], '.gw_', mm, '_beta_x_', rr, '_x_simagexcan.csv')
      df[[length(df) + 1]] = read.csv(tmp) %>% 
              mutate(idp_type = nn, phenotype = rr, model = mm)
    }
  }
}

df = do.call(rbind, df)
df_sb1 = df %>% mutate(zscore = p2z(pval, bhat))
# df = df %>% filter(is.na(stringr::str_match(IDP, 'PC')[, 1]))
```


Load genetic correlation results.

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/genetic_cor_2nd/', nn, '_2nd_x_gtex-gwas_x_', rr, '.ldsc_rg.log')
    tmp = load_ldsc_rg(tmp)
    tmp = tmp %>% select(p2, rg, p, z, h2_obs, h2_obs_se) %>% rename(IDP = p2, pval = p, zscore = z)
    df[[length(df) + 1]] = tmp %>% 
      mutate(idp_type = nn, phenotype = rr)
  }
}
df_cor1 = do.call(rbind, df)
```

## Load the psychiatric phenotypes

Load S-BrainXcan.

```{r}
traits = read.table('../misc_data/preprocess_psychiatric_traits/trait_list.txt')$V1
models = c('ridge', 'elastic_net')
tags = list(dmri = 'dmri.original.all_covar.w_pc', t1 = 't1.scaled.all_covar.w_pc')
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    for(mm in models) {
      tmp = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_psychiatric_2nd/', tags[[nn]], '.gw_', mm, '_beta_x_', rr, '_x_simagexcan.csv')
      df[[length(df) + 1]] = read.csv(tmp) %>% 
              mutate(idp_type = nn, phenotype = rr, model = mm)
    }
  }
}

df = do.call(rbind, df)
df_sb2 = df %>% mutate(zscore = p2z(pval, bhat))
# df = df %>% filter(is.na(stringr::str_match(IDP, 'PC')[, 1]))
```

Load genetic correlation results.

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/genetic_cor_2nd/', nn, '_2nd_x_psychiatric_x_', rr, '.ldsc_rg.log')
    tmp = load_ldsc_rg(tmp)
    tmp = tmp %>% select(p2, rg, p, z, h2_obs, h2_obs_se) %>% rename(IDP = p2, pval = p, zscore = z)
    df[[length(df) + 1]] = tmp %>% 
      mutate(idp_type = nn, phenotype = rr)
  }
}
df_cor2 = do.call(rbind, df)
```

## Load h2

```{r}
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    tmp = paste0('~/Desktop/tmp/ukb_idp/heritability_2nd_round/', tags[[nn]], '.tsv.gz')
    df[[length(df) + 1]] = read.table(tmp, sep = '\t', header = T) %>% mutate(idp_type = nn) %>% rename(IDP = phenotype)
  }
}
df_h2 = do.call(rbind, df)
```

## Combine 

```{r}
df_sb = rbind(
  df_sb1 %>% mutate(data = 'gtex-gwas'), 
  df_sb2 %>% mutate(data = 'psychiatric')
)
df_cor = rbind(
  df_cor1 %>% mutate(data = 'gtex-gwas'), 
  df_cor2 %>% mutate(data = 'psychiatric')
)
```

# S-BrainXcan vs genetic correlation

```{r, fig.width=6, fig.height=6}
tmp = inner_join(df_sb, df_cor, by = c('phenotype', 'idp_type', 'IDP', 'data'), suffix = c('.SB', '.RG'))
tmp %>% ggplot() + geom_point(aes(x = zscore.RG, y = zscore.SB), alpha = 0.2) + facet_grid(model~data) + th2 + geom_hline(yintercept = 0, color = 'gray') + geom_vline(xintercept = 0, color = 'gray')
```

```{r, fig.width=6, fig.height=6}
# tmp = inner_join(df_sb, df_cor, by = c('phenotype', 'idp_type', 'IDP', 'data'), suffix = c('.SB', '.RG'))
tmp %>% filter(pval.SB < 1e-2) %>% ggplot() + geom_point(aes(x = rg, y = bhat), alpha = 0.2) + facet_grid(model~data) + th2 + 
  scale_y_continuous(lim = c(-1, 1)) +
  geom_hline(yintercept = 0, color = 'gray') + geom_vline(xintercept = 0, color = 'gray')
```

# GCTA h2 vs LDSC h2

```{r}
tmp = inner_join(df_h2, df_cor, by = c('idp_type', 'IDP'), suffix = c('.GCTA', '.LDSC'))
tmp = tmp %>% group_by(IDP, idp_type) %>% 
  summarize(
    h2.LDSC = mean(h2_obs), h2_se.LDSC = max(h2_obs_se), 
    h2.GCTA = h2[1], h2_se.GCTA = h2_SE[1]
  ) %>% ungroup()
tmp %>% ggplot() + geom_point(aes(x = h2.GCTA, y = h2.LDSC)) + 
  geom_errorbar(aes(x = h2.GCTA, ymin = h2.LDSC - 1.96 * h2_se.LDSC, ymax = h2.LDSC + 1.96 * h2_se.LDSC), alpha = 0.2) + 
  geom_errorbarh(aes(y = h2.LDSC, xmin = h2.GCTA - 1.96 * h2_se.GCTA, xmax = h2.GCTA + 1.96 * h2_se.GCTA), alpha = 0.2) +
  facet_wrap(~idp_type) + th2 + geom_abline(slope = 1, intercept = 0, color = 'gray')
```
