---
title: "Round 2: S-BrainXcan results of psychiatric phenotypes (select scaled T1 and original dMRI)"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://raw.githubusercontent.com/liangyy/misc-tools/master/plot_tool/plot_tools.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(ggrepel)
library(data.table)
options(datatable.fread.datatable = F)
source('rlib.R')
library(VennDiagram)
library(ggpubr)
```

# About

This is the selected psychiatric phenotypes S-BrainXcan preliminary results of the 2nd round of the BrainXcan. 
In particular, we focus on 

* `scaled_t1_all_covar_w_pc`
* `orignal_dmri_all_covar_w_pc`

# Load data

```{r}
traits = read.table('../misc_data/preprocess_psychiatric_traits/trait_list.txt')$V1
models = c('ridge', 'elastic_net')
tags = list(dmri = 'dmri.original.all_covar.w_pc', t1 = 't1.scaled.all_covar.w_pc')
df = list()
for(rr in traits) {
  for(nn in names(tags)) {
    for(mm in models) {
      tmp = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results_psychiatric_2nd/', tags[[nn]], '.gw_', mm, '_beta_x_', rr, '_x_simagexcan.csv')
      df[[length(df) + 1]] = read.csv(tmp) %>% 
              mutate(idp_type = nn, phenotype = rr, model = mm)
    }
  }
}

df = do.call(rbind, df)
df = df %>% mutate(zscore = p2z(pval, bhat))
# df = df %>% filter(is.na(stringr::str_match(IDP, 'PC')[, 1]))
```


Load IDP annotations. 

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
dmri = readRDS('../misc_data/download_some_matching_files/annot_dmri_idps.rds')
dmri = dmri %>% mutate(IDP = paste0('IDP-', FieldID))
idps = c('t1', 'dmri')
df_annot = rbind(
  t1 %>% select(IDP, Field) %>% mutate(idp_type = 't1'),
  dmri %>% select(IDP, Field) %>% mutate(idp_type = 'dmri')
)
```

# Overview of the results

```{r, fig.height=4, fig.width=5}
p1 = qqplot_by_group(df$pval, paste(df$model, df$idp_type)) + th + ggtitle('QQ-plot: Psychiatric phenotypes') 
p1
```


```{r, fig.height=4, fig.width=8}
p1 = df %>% reshape2::dcast(IDP + phenotype + idp_type ~ model, value.var = 'pip') %>% ggplot() +
  geom_point(aes(x = ridge, y = elastic_net), alpha = 0.05) +
  th2 + ggtitle('PIP: Psychiatric phenotypes') + facet_wrap(~idp_type)
p1 
```



# Selecting putative signals

Meet both criteria:

* Bonferroni significant at $\alpha = 0.01$ (we used stringent cutoff since there is some sort of inflation).
* Within 95% credible set.

```{r}
df_signif = df %>% group_by(idp_type, model) %>% mutate(p_adj = pval * n()) %>% ungroup() %>%
  filter(cs95 > 0 & p_adj < 0.01)

df_signif %>% group_by(idp_type, model) %>% summarize(nsig = n()) %>% pander::pander(caption = 'The number of putative signals extracted from Psychiatric phenotypes')
```

Save putative signals to TSV.

```{r}
for(dd in c('dmri', 't1')) {
  sub = df_signif %>% filter(idp_type == dd) %>% select(IDP, phenotype) %>% distinct()
  sub$pheno2 = sub$phenotype
  colnames(sub) = c('idp', 'pheno', 'pheno_code')
  sub = sub[, c('pheno', 'pheno_code', 'idp')]
  write.table(sub, paste0('../misc_data/round_2_selected.psychiatric.', dd, '.signif.tsv'), quo = F, col = T, row = F, sep = '\t')
}
```

# Mendelian randomization results

```{r}
collector_mr = list()
collector_data = list()
idps = c('t1', 'dmri') 
df_signif_out = df_signif %>% select(idp_type, IDP, phenotype) %>% distinct()
for(dd in idps) {
  sub = df_signif_out %>% filter(idp_type == dd) %>% select(-idp_type)
  for(i in 1 : nrow(sub)) {
    tmp = readRDS(paste0('~/Desktop/tmp/ukb_idp/mr_psychiatric_2nd/MR_local.psychiatric_2nd_', dd, '.', sub$IDP[i], '_x_', sub$phenotype[i], '.rds'))
    if(sum(is.na(tmp$idp2pheno$mr)) == 1 | sum(is.na(tmp$pheno2idp$mr)) == 1) {
      next
    }
    n_idp2pheno = sum(tmp$idp2pheno$mr$pval < 0.05) 
    n_pheno2idp = sum(tmp$pheno2idp$mr$pval < 0.05) 
    if(n_idp2pheno > 2 | n_pheno2idp > 2) {
      collector_mr[[length(collector_mr) + 1]] = rbind(
        tmp$idp2pheno$mr %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd),
        tmp$pheno2idp$mr %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd)
      )
      collector_data[[length(collector_data) + 1]] = rbind(
        tmp$idp2pheno$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'idp2pheno', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd),
        tmp$pheno2idp$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'pheno2idp', pheno = sub$phenotype[i], IDP = sub$IDP[i], is_forward = n_idp2pheno > 3, is_pass = n_idp2pheno > 3 | n_pheno2idp > 3, idp_type = dd)
      )
    } 
  }
}

df_data = do.call(rbind, collector_data) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr = do.call(rbind, collector_mr) %>% left_join(df_annot, by = c('IDP', 'idp_type'))
df_mr$Field[ is.na(df_mr$Field) ] = paste(df_mr$idp_type[ is.na(df_mr$Field) ], df_mr$IDP[ is.na(df_mr$Field) ])
df_data$Field[ is.na(df_data$Field) ] = paste(df_data$idp_type[ is.na(df_data$Field) ], df_data$IDP[ is.na(df_data$Field) ])
```


```{r, fig.width=15, fig.height=5}
df_all_signif = df
plot_pair = function(pheno_, idp_, type_) {
  mydat = df_data %>% filter(pheno == pheno_, IDP == idp_, idp_type == type_)
  mymr = df_mr %>% filter(pheno == pheno_, IDP == idp_, idp_type == type_)
  idp_name = mydat$Field[1]
  gwas_name = mydat$pheno[1]
  idp_id = mydat$IDP[1]
  this_result = df_all_signif %>% filter(IDP == idp_id, phenotype == gwas_name, idp_type == type_)
  plist = list()
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'idp2pheno') %>% ggplot() + 
    geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + 
    geom_vline(xintercept = 0, color = 'lightgray') + 
    geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(idp_name, ' -> \n ', gwas_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = bhat, intercept = 0, color = model))
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'pheno2idp') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(gwas_name, ' -> \n ', idp_name)) +
    theme(legend.position = 'none') # +
  # geom_abline(data = this_result, aes(slope = 1 / bhat, intercept = 0, color = model))
  # plist[[length(plist) + 1]] = gridExtra::tableGrob(df_mr %>% select(method, nsnp, b, pval, direction))
  this_result2 = this_result %>% mutate(method = paste0('imgx_', model), nsnp = paste0('pip = ', signif(pip, digits = 2), '\n cs = ', cs95), b = bhat, direction = NA) 
  mymr = rbind(mymr %>% select(method, nsnp, b, pval, direction), this_result2 %>% select(method, nsnp, b, pval, direction))
  qq = ggarrange(plist[[1]], plist[[2]], ggtexttable(format(mymr, digits = 3)), ncol = 3)
  qq
}
```


```{r, fig.width=15, fig.height=5}
df_tmp = df_mr %>% select(pheno, IDP, idp_type) %>% distinct()  # %>% filter(is_pass == T) 
for(i in 1 : nrow(df_tmp)) {
  qq = plot_pair(df_tmp$pheno[i], df_tmp$IDP[i], df_tmp$idp_type[i])
  print(qq)
}

```

