---
title: "Round 2: heritability"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(ggrepel)
library(data.table)
options(datatable.fread.datatable = F)
pheno_interest = c('weekly_alcohol', 'recurrent_depressive_disorder', 'parent_depression', 'parent_AD', 'handedness', 'daily_coffee', 'daily_cigarettes')
pheno_bcc = c('wbc', 'rbc', 'platelet', 'lymphocyte', 'monocyte', 'neutrophil', 'eosinophil', 'basophil')
pheno_ht = c('dbp', 'sbp') # , 'ht', 'hb', 'mcv', 'mch', 'mchc')
pheno_rand = paste0('random_pheno_', 0 : 9)
pheno_sig_dmri = paste0('single_sig_dmri_', 0 : 9)
pheno_sig_orig = paste0('single_sig_orig_', 0 : 9)
source('rlib.R')
```

# About

This is the heritabiltiy preliminary results of the 2nd round of the BrainXcan. 
In particular, we run and compare various ways to do phenotype pre-processing.

* `orignal_t1_all_covar_no_pc`
* `orignal_t1_non_idp_covar_no_pc`
* `orignal_t1_all_covar_w_pc`
* `orignal_t1_non_idp_covar_w_pc`
* `scaled_t1_all_covar_no_pc`
* `scaled_t1_non_idp_covar_no_pc`
* `scaled_t1_all_covar_w_pc`
* `scaled_t1_non_idp_covar_w_pc`
* `regress_t1_all_covar_no_pc`
* `regress_t1_non_idp_covar_no_pc`
* `regress_t1_all_covar_w_pc`
* `regress_t1_non_idp_covar_w_pc`
* `orignal_dmri_all_covar_no_pc`
* `orignal_dmri_non_idp_covar_no_pc`
* `orignal_dmri_all_covar_w_pc`
* `orignal_dmri_non_idp_covar_w_pc`
* `regress_dmri_all_covar_no_pc`
* `regress_dmri_non_idp_covar_no_pc`
* `regress_dmri_all_covar_w_pc`
* `regress_dmri_non_idp_covar_w_pc`

# Load data

```{r}
types = c('scaled', 'regress', 'original')
idps = c('dmri', 't1')
covars = c('all_covar', 'non_idp_covar')
pcs = c('no_pc', 'w_pc')
df = list()
for(tt in types) {
  for(ii in idps) {
    for(cc in covars) {
      for(pp in pcs) {
        if(ii == 'dmri' & tt == 'scaled') {
          next
        }
        tmp = paste0('~/Desktop/tmp/ukb_idp/heritability_2nd_round/', ii, '.', tt, '.', cc, '.', pp, '.tsv.gz')
        df[[length(df) + 1]] = read.table(tmp, sep = '\t', header = T) %>% mutate(scale_type = tt, idp_type = ii, pc_type = pp, covar_type = cc)
      }
    }
  }
}
df = do.call(rbind, df)

df_old = list()
for(ii in idps) {
  tmp = paste0('~/Desktop/tmp/ukb_idp/heritability/pyemma.idp_', ii, '.tsv.gz')
  df_old[[length(df_old) + 1]] = read.table(tmp, sep = '\t', header = T) %>% mutate(idp_type = ii)
}
df_old[[length(df_old) + 1]] = read.table('~/Desktop/tmp/ukb_idp/heritability/pyemma.idp_original.tsv.gz', sep = '\t', header = T) %>% mutate(idp_type = 'original')
df_old = do.call(rbind, df_old)
```

<!-- ```{r, fig.width=3, fig.height=3} -->
<!-- df_old %>% ggplot() + geom_histogram(aes(x = h2)) + facet_wrap( ~ idp_type, ncol = 1, scales = 'free_y') + ggtitle('Previous results') -->
<!-- ``` -->

# T1

```{r, fig.width=12, fig.height=4}
p1 = df %>% filter(idp_type == 't1', pc_type == 'no_pc') %>% ggplot() + geom_histogram(aes(x = h2)) + facet_grid(scale_type ~ covar_type) + ggtitle('T1: no PC adjustment') + th2
p2 = df %>% filter(idp_type == 't1', pc_type == 'w_pc') %>% ggplot() + geom_histogram(aes(x = h2)) + facet_grid(scale_type ~ covar_type) + ggtitle('T1: with PC adjustment') + th2
p1 + p2
```

## Old vs new 

The `regress.all_covar.w_pc` matches best to the previous result.
They differ by whether to inverse normalize the residual of PCs or not. 

```{r, fig.width=4, fig.height=4}
tmp = inner_join(df_old %>% filter(idp_type == 't1'), df %>% filter(idp_type == 't1', covar_type == 'all_covar', scale_type == 'regress', pc_type == 'w_pc'), by = 'phenotype', suffix = c('.old', '.new'))
tmp %>% ggplot() + geom_point(aes(x = h2.old, y = h2.new), alpha = 0.5) + geom_text_repel(data = tmp %>% filter(abs(h2.new - h2.old) > 0.05), aes(x = h2.old, y = h2.new, label = phenotype), size = 3) + th + coord_equal() + geom_abline(slope = 1, intercept = 0)
```

## Original vs regress

We can see the bimodal shape when the in regress/scaled. 
To look into the source, we compare regress and original for no PC adjustment.

```{r, fig.width=4, fig.height=4}
tmp = df %>% filter(idp_type == 't1', covar_type == 'all_covar', pc_type == 'no_pc') %>% reshape2::dcast(phenotype ~ scale_type, value.var = 'h2')
tmp %>% ggplot() + geom_point(aes(x = original, y = regress), alpha = 0.5) + geom_text_repel(data = tmp %>% filter(abs(original - regress) < 0.001), aes(x = original, y = regress, label = phenotype), size = 3) + th + coord_equal() + geom_abline(slope = 1, intercept = 0)
```

The observation is that some IDPs have reduced h2 when adjusting for covariates but others do not. The same observation appears when comparing the `orignal.all_covar.no_pc` with the previous h2 calculated on the Owen's original data (no PC adjustment).

```{r, fig.width=4, fig.height=4}
tmp = inner_join(df_old %>% filter(idp_type == 'original'), df %>% filter(idp_type == 't1', covar_type == 'all_covar', scale_type == 'original', pc_type == 'no_pc'), by = 'phenotype', suffix = c('.old', '.new'))
tmp %>% ggplot() + geom_point(aes(x = h2.new, y = h2.old), alpha = 0.5) + 
  geom_text_repel(data = tmp %>% filter(abs(h2.new - h2.old) < 0.001), aes(x = h2.old, y = h2.new, label = phenotype), size = 3) +
  th + coord_equal() + geom_abline(slope = 1, intercept = 0)
```


## scaled vs regress

```{r, fig.width=6, fig.height=3}
tmp = df %>% filter(idp_type == 't1', covar_type == 'all_covar', pc_type == 'no_pc') %>% reshape2::dcast(phenotype ~ scale_type, value.var = 'h2')
p1 = tmp %>% ggplot() + geom_point(aes(x = regress, y = scaled), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(scaled - regress) < 0.001), aes(x = regress, y = scaled, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) + ggtitle('no PC')
tmp = df %>% filter(idp_type == 't1', covar_type == 'all_covar', pc_type == 'w_pc') %>% reshape2::dcast(phenotype ~ scale_type, value.var = 'h2')
p2 = tmp %>% ggplot() + geom_point(aes(x = regress, y = scaled), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(scaled - regress) < 0.001), aes(x = regress, y = scaled, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) + ggtitle('with PC')
p1 + p2
```

Overall they agree and scaled ones tend to have slightly higher h2.


# With/without PC

For T1, we may or may not want to use PC adjustment but we do want to use "scaled" scale.

```{r, fig.width=3, fig.height=3}
tmp = df %>% filter(idp_type == 't1', covar_type == 'all_covar', scale_type == 'scaled') %>% reshape2::dcast(phenotype ~ pc_type, value.var = 'h2')
p1 = tmp %>% ggplot() + geom_point(aes(x = no_pc, y = w_pc), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(original - regress) < 0.001), aes(x = original, y = regress, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) 
p1
```

# dMRI

```{r, fig.width=12, fig.height=4}
p1 = df %>% filter(idp_type == 'dmri', pc_type == 'no_pc') %>% ggplot() + geom_histogram(aes(x = h2)) + facet_grid(scale_type ~ covar_type) + ggtitle('dMRI: no PC adjustment') + th2
p2 = df %>% filter(idp_type == 'dmri', pc_type == 'w_pc') %>% ggplot() + geom_histogram(aes(x = h2)) + facet_grid(scale_type ~ covar_type) + ggtitle('dMRI: with PC adjustment') + th2
p1 + p2
```

Overall, they agree well.

## Old vs new 

The `regress.all_covar.w_pc` matches best to the previous result.
They differ by whether to inverse normalize the residual of PCs or not. 

```{r, fig.width=4, fig.height=4}
tmp = inner_join(df_old %>% filter(idp_type == 'dmri'), df %>% filter(idp_type == 'dmri', covar_type == 'all_covar', scale_type == 'regress', pc_type == 'w_pc'), by = 'phenotype', suffix = c('.old', '.new'))
tmp %>% ggplot() + geom_point(aes(x = h2.old, y = h2.new), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(h2.new - h2.old) > 0.05), aes(x = h2.old, y = h2.new, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0)
```


## Original vs regress

Since we think dMRI should be used in the original scale, we compare original and regress scaling. 

```{r, fig.width=6, fig.height=3}
tmp = df %>% filter(idp_type == 'dmri', covar_type == 'all_covar', pc_type == 'no_pc') %>% reshape2::dcast(phenotype ~ scale_type, value.var = 'h2')
p1 = tmp %>% ggplot() + geom_point(aes(x = original, y = regress), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(original - regress) < 0.001), aes(x = original, y = regress, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) + ggtitle('no PC')
tmp = df %>% filter(idp_type == 'dmri', covar_type == 'all_covar', pc_type == 'w_pc') %>% reshape2::dcast(phenotype ~ scale_type, value.var = 'h2')
p2 = tmp %>% ggplot() + geom_point(aes(x = original, y = regress), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(original - regress) < 0.001), aes(x = original, y = regress, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) + ggtitle('with PC')
p1 + p2
```

They have very similar h2.

# With/without PC

For dMRI, we may want to use PC adjustment in original scale.

```{r, fig.width=3, fig.height=3}
tmp = df %>% filter(idp_type == 'dmri', covar_type == 'all_covar', scale_type == 'original') %>% reshape2::dcast(phenotype ~ pc_type, value.var = 'h2')
p1 = tmp %>% ggplot() + geom_point(aes(x = no_pc, y = w_pc), alpha = 0.5) + 
  # geom_text_repel(data = tmp %>% filter(abs(original - regress) < 0.001), aes(x = original, y = regress, label = phenotype), size = 3) + 
  th + coord_equal() + geom_abline(slope = 1, intercept = 0) 
p1
```
