---
title: "Summary ImageXcan GTEx-GWAS run"
# author: Yanyu Liang
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup}
rm(list = ls())
library(ggplot2)
theme_set(theme_bw(base_size = 15))
source('https://gist.githubusercontent.com/liangyy/43912b3ecab5d10c89f9d4b2669871c9/raw/3ca651cfa53ffccb8422f432561138a46e93710f/my_ggplot_theme.R')
source('https://raw.githubusercontent.com/liangyy/misc-tools/master/plot_tool/plot_tools.R')
library(dplyr)
options(stringsAsFactors = F)
library(patchwork)
library(data.table)
options(datatable.fread.datatable = F)
source('rlib.R')
library(ggpubr)
library(pander)
panderOptions("table.split.table", Inf)

```

# About

Here I ran S-ImageXcan on 114 GTEx-GWAS phenotypes.

# Load results

```{r}
pheno_list = read.table('../submission/simagexcan/gtex_gwas_list.txt', header = F)$V1
df = list()
for(idp_type in c('dmri', 't1')) {
  for(p in pheno_list) {
    fn = paste0('~/Desktop/tmp/ukb_idp/simagexcan/results/', idp_type, '_gtex_gwas_x_', p, '_x_simagexcan.csv')
    if(file.exists(fn)) {
      tmp = read.csv(fn)
      df[[length(df) + 1]] = tmp %>% mutate(phenotype = p, idp_type = idp_type)
    }
  }
}

df = do.call(rbind, df)
```

Load dMRI and T1 IDP annotations.

```{r}
t1 = readRDS('../misc_data/process_t1/t1_meta.rds')
t1 = t1 %>% mutate(IDP = paste0('IDP-', FieldID))
dmri = readRDS('../misc_data/download_some_matching_files/annot_dmri_idps.rds')
dmri = dmri %>% mutate(IDP = paste0('IDP-', FieldID))
idps = c('t1', 'dmri')
df_annot = rbind(
  t1 %>% select(IDP, Field),
  dmri %>% select(IDP, Field)
)
```

# QQ-plot

```{r}
qqplot_by_group(df$pval, group = df$idp_type) + th
```

# PIP

```{r, fig.width=8, fig.height=6}
df %>% ggplot() + geom_jitter(aes(x = phenotype, y = pip, color = factor(cs95)), width = 0.3, height = 0, alpha = 0.2) + theme(axis.text.x = element_blank()) + th2 + facet_wrap(~idp_type, ncol = 1)
```

# Bonferroni significant

```{r}
df = df %>% mutate(p_adj = pval * n())
bonferroni_cutoff = 0.01
message('There are ', df %>% filter(p_adj < bonferroni_cutoff) %>% nrow, ' IDP-phenotype pairs passed Bonferroni significance cutoff at ', bonferroni_cutoff)
```

# 95% Credible set

```{r}
message('There are ', df %>% filter(cs95 > 0) %>% nrow, ' IDP-phenotype pairs within 95% credible set')
```

# Add open GWAS phenotype

Here we focus on the phenotypes with at least one hit within 95% credible set.


```{r}
df_cs = df %>% filter(cs95 > 0)
phenos = c('Astle_et_al_2016_Eosinophil_counts', 'Astle_et_al_2016_Granulocyte_count', 'Astle_et_al_2016_High_light_scatter_reticulocyte_count', 'Astle_et_al_2016_Lymphocyte_counts', 'Astle_et_al_2016_Monocyte_count', 'Astle_et_al_2016_Myeloid_white_cell_count', 'Astle_et_al_2016_Neutrophil_count', 'Astle_et_al_2016_Platelet_count', 'Astle_et_al_2016_Red_blood_cell_count', 'Astle_et_al_2016_Reticulocyte_count', 'Astle_et_al_2016_Sum_basophil_neutrophil_counts', 'Astle_et_al_2016_Sum_eosinophil_basophil_counts', 'Astle_et_al_2016_Sum_neutrophil_eosinophil_counts', 'Astle_et_al_2016_White_blood_cell_count', 'EGG_BL', 'EGG_BW3_EUR', 'GIANT_2015_BMI_EUR', 'GIANT_HEIGHT', 'IBD.EUR.Crohns_Disease', 'RA_OKADA_TRANS_ETHNIC', 'SSGAC_Depressive_Symptoms', 'SSGAC_Education_Years_Pooled', 'UKB_1160_Sleep_duration', 'UKB_1180_Morning_or_evening_person_chronotype', 'UKB_1807_Fathers_age_at_death', 'UKB_20002_1065_self_reported_hypertension', 'UKB_20002_1111_self_reported_asthma', 'UKB_20002_1225_self_reported_hyperthyroidism_or_thyrotoxicosis', 'UKB_20002_1309_self_reported_osteoporosis', 'UKB_20002_1453_self_reported_psoriasis', 'UKB_20002_1473_self_reported_high_cholesterol', 'UKB_20016_Fluid_intelligence_score', 'UKB_20022_Birth_weight', 'UKB_20127_Neuroticism_score', 'UKB_21001_Body_mass_index_BMI', 'UKB_23099_Body_fat_percentage', 'UKB_2395_4_Hair_or_balding_pattern_Pattern_4', 'UKB_3526_Mothers_age_at_death', 'UKB_50_Standing_height', 'UKB_6150_1_Vascular_or_heart_problems_diagnosed_by_doctor_Heart_attack', 'UKB_6152_8_diagnosed_by_doctor_Asthma', 'UKB_6152_9_diagnosed_by_doctor_Hayfever_allergic_rhinitis_or_eczema', 'pgc.scz2', 'BCAC_Overall_BreastCancer_EUR', 'GIANT_2017_BMI_Active_EUR', 'GIANT_HIP_Combined_EUR', 'GIANT_WC_Combined_EUR', 'IMMUNOBASE_Multiple_sclerosis_hg19', 'UKB_1200_Sleeplessness_or_insomnia', 'UKB_20002_1154_self_reported_irritable_bowel_syndrome', 'UKB_20002_1226_self_reported_hypothyroidism_or_myxoedema', 'UKB_20002_1261_self_reported_multiple_sclerosis', 'UKB_20002_1463_self_reported_ulcerative_colitis', 'UKB_2395_2_Hair_or_balding_pattern_Pattern_2', 'ENIGMA_Intracraneal_Volume', 'GLGC_Mc_TG')
gwas_code = c('ebi-a-GCST004606', 'ebi-a-GCST004614', 'ukb-d-30300_irnt', 'ebi-a-GCST004627', 'ukb-d-30130_irnt', 'ebi-a-GCST004626', 'ebi-a-GCST004629', 'ukb-d-30080_irnt', 'ebi-a-GCST004601', 'ukb-d-30250_irnt', 'ebi-a-GCST004620', 'ebi-a-GCST004621', 'ebi-a-GCST004613', 'ieu-b-30', 'ieu-a-29', 'ukb-b-13378', 'ukb-b-19953', 'ieu-a-89', 'ukb-a-552', 'ukb-b-11874', 'ieu-a-1000', 'ukb-b-6134', 'ukb-b-4424', 'ukb-b-4956', 'ukb-b-11303', 'ukb-b-12493', 'ukb-b-20208', 'ukb-b-20289', 'ukb-b-17796', 'ukb-b-10537', 'ukb-b-10912', 'ukb-b-5238', 'ukb-b-13378', 'ukb-b-4630', 'ukb-b-19953', 'ukb-b-8909', 'ukb-a-303', 'ukb-b-12687', 'ukb-b-10787', 'ukb-b-11590', 'ukb-b-20208', 'ukb-b-17241', 'ieu-a-22', 'ieu-a-1126', 'ukb-b-19953', 'ukb-b-15590', 'ukb-b-9405', 'ieu-b-18', 'ukb-b-3957', 'ukb-b-2592', 'ukb-b-19732', 'ieu-b-18', 'ukb-b-7584', 'ukb-a-301', 'ieu-a-1041', 'ieu-a-302')
df_ogwas = data.frame(phenotype = phenos, gwas_code = gwas_code)
df_signif = inner_join(df_ogwas, df_cs, by = c('phenotype' = 'phenotype'))
df_signif = df_signif[ !is.na(df_signif$gwas_code), ]
df_signif_t1 = df_signif %>% filter(idp_type == 't1') %>% select(phenotype, gwas_code, IDP) %>% distinct_all
colnames(df_signif_t1) = c('pheno', 'pheno_code', 'idp')
df_signif_dmri = df_signif %>% filter(idp_type == 'dmri') %>% select(phenotype, gwas_code, IDP) %>% distinct_all
colnames(df_signif_dmri) = c('pheno', 'pheno_code', 'idp')
write.table(df_signif_t1, '../misc_data/simagexcan_gtex_gwas.t1.signif.tsv', quo = F, col = T, row = F, sep = '\t')
write.table(df_signif_dmri, '../misc_data/simagexcan_gtex_gwas.dmri.signif.tsv', quo = F, col = T, row = F, sep = '\t')
```

# MR results

```{r}
collector_mr = list()
collector_data = list()
for(i in 1 : nrow(df_signif)) {
  tmp = readRDS(paste0('~/Desktop/tmp/ukb_idp/mr_gtex_gwas/MR.gtex_gwas_', df_signif$idp_type[i], '.', df_signif$IDP[i], '_x_', df_signif$pheno[i], '.rds'))
  if(sum(is.na(tmp$idp2pheno$mr)) == 1 | sum(is.na(tmp$pheno2idp$mr)) == 1) {
    next
  }
  if(sum(tmp$idp2pheno$mr$pval < 0.05) > 2 | sum(tmp$pheno2idp$mr$pval < 0.05) > 2) {
    collector_mr[[length(collector_mr) + 1]] = rbind(
      tmp$idp2pheno$mr %>% mutate(direction = 'idp2pheno', pheno = df_signif$pheno[i], IDP = df_signif$IDP[i]),
      tmp$pheno2idp$mr %>% mutate(direction = 'pheno2idp', pheno = df_signif$pheno[i], IDP = df_signif$IDP[i])
    )
    collector_data[[length(collector_data) + 1]] = rbind(
      tmp$idp2pheno$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'idp2pheno', pheno = df_signif$pheno[i], IDP = df_signif$IDP[i]),
      tmp$pheno2idp$data %>% select(beta.exposure, beta.outcome, pval.exposure) %>% mutate(direction = 'pheno2idp', pheno = df_signif$pheno[i], IDP = df_signif$IDP[i])
    )
  }
}
df_data = do.call(rbind, collector_data) %>% inner_join(df_annot, by = 'IDP')
df_mr = do.call(rbind, collector_mr) %>% inner_join(df_annot, by = 'IDP')
```

The selected MR pairs.

```{r}
# df_mr %>% select(pheno, IDP) %>% distinct() %>% pander::pander()
df_tmp = df_mr %>% select(pheno, IDP) %>% distinct()
```

Show the plot.

```{r, fig.width=15, fig.height=5}
for(i in 1 : nrow(df_tmp)) {
  mydat = df_data %>% filter(pheno == df_tmp$pheno[i], IDP == df_tmp$IDP[i])
  mymr = df_mr %>% filter(pheno == df_tmp$pheno[i], IDP == df_tmp$IDP[i])
  idp_name = mydat$Field[1]
  gwas_name = mydat$pheno[1]
  plist = list()
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'idp2pheno') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(idp_name, ' -> \n ', gwas_name)) +
    theme(legend.position = 'none')
  plist[[length(plist) + 1]] = mydat %>% filter(direction == 'pheno2idp') %>% ggplot() + geom_point(aes(x = beta.exposure, y = beta.outcome, size = -log10(pval.exposure))) + geom_vline(xintercept = 0, color = 'lightgray') + geom_hline(yintercept = 0, color = 'lightgray') + th + ggtitle(paste0(gwas_name, ' -> \n ', idp_name)) +
    theme(legend.position = 'none')
  # plist[[length(plist) + 1]] = gridExtra::tableGrob(df_mr %>% select(method, nsnp, b, pval, direction))
  qq = ggarrange(plist[[1]], plist[[2]], ggtexttable(format(mymr %>% select(method, nsnp, b, pval, direction), digits = 3)), ncol = 3)
  print(qq)
}

```


# List the pairs with at least 3 MR tests p-value < 0.05

```{r}
df_mr %>% filter(pval < 0.05) %>% group_by(direction, pheno, Field) %>% summarize(n_sig_pval = n()) %>% filter(n_sig_pval >= 3) %>% pander()
```

# Side-by-side with ImageXcan on blood cell counts

```{r}
p2z = function(p, b) {
  sign(b) * qnorm(p / 2)
}

df_bcc = data.frame(
  id = c('wbc', 'rbc', 'platelet', 'lymphocyte', 'monocyte', 'neutrophil', 'eosinophil'),
  gwas = c('Astle_et_al_2016_White_blood_cell_count', 'Astle_et_al_2016_Red_blood_cell_count', 'Astle_et_al_2016_Platelet_count', 'Astle_et_al_2016_Lymphocyte_counts', 'Astle_et_al_2016_Monocyte_count', 'Astle_et_al_2016_Neutrophil_count', 'Astle_et_al_2016_Eosinophil_counts')
)
df_t1_indiv = inner_join(
  read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.linear.t1.csv'),
  read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.susie.t1.csv'),
  by = c('phenotype', 'IDP')
) %>% mutate(idp_type = 't1') %>% filter(phenotype %in% df_bcc$id)
df_dmri_indiv = inner_join(
  read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.linear.dmri.csv'),
  read.csv('~/Desktop/tmp/ukb_idp/data/imagexcan_round_1.susie.dmri.csv'),
  by = c('phenotype', 'IDP')
) %>% mutate(idp_type = 'dmri') %>% filter(phenotype %in% df_bcc$id)
df_indiv = rbind(
  df_t1_indiv, df_dmri_indiv
)
df_ss = df %>% filter(phenotype %in% df_bcc$gwas) %>% inner_join(df_bcc, by = c('phenotype' = 'gwas'))
tmp = inner_join(df_indiv, df_ss, by = c('phenotype' = 'id', 'IDP', 'idp_type'), suffix = c('.indiv', '.ss'))
tmp %>% 
  # filter(pval.indiv < 0.05, pval.ss < 0.05) %>% 
  group_by(phenotype) %>% 
  filter(abs(bhat.ss) < quantile(abs(bhat.ss), probs = 0.95), abs(bhat.indiv) < quantile(abs(bhat.indiv), probs = 0.975)) %>% 
  ggplot() + geom_point(aes(x = bhat.indiv, y = bhat.ss), alpha = 0.2) + facet_wrap(~phenotype, scales = 'free') + th2
tmp %>% 
  # filter(pval.indiv < 0.05, pval.ss < 0.05) %>% 
  group_by(phenotype) %>% 
  # filter(abs(bhat.ss) < quantile(abs(bhat.ss), probs = 0.95), abs(bhat.indiv) < quantile(abs(bhat.indiv), probs = 0.975)) %>% 
  mutate(z.indiv = p2z(pval.indiv, bhat.indiv), z.ss = p2z(pval.ss, bhat.ss)) %>% 
  ggplot() + geom_point(aes(x = z.indiv, y = z.ss), alpha = 0.2) + facet_wrap(~phenotype) + th2 + coord_equal() + geom_abline(slope = -1, intercept = 0, color = 'red')
tmp %>% 
  # filter(pval.indiv < 0.05, pval.ss < 0.05) %>% 
  group_by(phenotype) %>% 
  # filter(abs(bhat.ss) < quantile(abs(bhat.ss), probs = 0.95), abs(bhat.indiv) < quantile(abs(bhat.indiv), probs = 0.975)) %>% 
  # mutate(z.indiv = p2z(pval.indiv, bhat.indiv), z.ss = p2z(pval.ss, bhat.ss)) %>% 
  ggplot() + geom_point(aes(x = pip.indiv, y = pip.ss), alpha = 0.2) + facet_wrap(~phenotype) + th2 + coord_equal() 
```