---
title: "Brief summary of IWAS results"
author: "Yanyu Liang"
date: "5/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(stringsAsFactors = F)
library(ggplot2)
theme_set(theme_classic(base_size = 12))
library(dplyr)
library(pander)
panderOptions('table.split.table', Inf)
```

# About

Here I summarize some of the results shared by IWAS paper [https://www.sciencedirect.com/science/article/pii/S1053811920308338](https://www.sciencedirect.com/science/article/pii/S1053811920308338).

I focused on structural and diffusion IDPs from UKB part (supplementary table).

# Load data

```{r}
kk = rbind(
  read.csv('../iwas_s1_structural.csv')[-1, ],
  read.csv('../iwas_s1_diffusion.csv')[-1, ]
)
ff = rbind(
  read.csv('../iwas_s1_structural.csv')[1, ] %>% mutate(type = 'structural'),
  read.csv('../iwas_s1_diffusion.csv')[1, ] %>% mutate(type = 'diffusion')
)
kk %>% ggplot() + geom_point(aes(x = beta_MVIWAS, y = beta_MVIWAS_Egger)) + coord_equal()
kk %>% ggplot() + geom_point(aes(x = -log(p_MVIWAS), y = -log(p_MVIWAS_Egger))) + coord_equal()
ff[, c(-2:-7, -11)] %>% pander(caption = 'Mu in Egger')
```

# A brief summary

Differences between BrainXcan and IWAS:

* They did not do decomposition (PCA for obtaining region-specific and common factors).
* They proposed MV-IWAS which fit multiple IDPs jointly.
* They also proposed MV-IWAS-Egger which considers the direct effect of SNPs.
    - But the Egger version does not change the result much (see Figures above).
    
Pitfalls of IWAS:

* They did not use up-to-date UKB IDP GWAS.
* The joint model is sensitive to LD.
* The MV-IWAS was done for significant IDPs which is subject to selective bias and they do not account for this bias. 



