---
output:
  html_document: 
    self_contained: yes
    toc: true
    toc_float: true
    toc_collapsed: true
    number_sections: true
    theme: cosmo
params:
  phenotype_name: "phenotype of interest"
  color_code_yaml: NA
  input_prefix: NA
  rlib: NA
  ntop: 10
  bxcan_pval_col: 'pval'
title: "`r paste0('BrainXcan results for ', params$phenotype_name)`"
---


```{r setup, include=FALSE}
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(pander))
options(stringsAsFactors = FALSE)
theme_set(theme_classic(base_size = 12))
panderOptions('table.split.table', Inf)
source('/Users/yanyuluchicago/Documents/repo/GitHub/brainxcan/brainxcan/vis/report_helper.R')
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(echo = FALSE)
```

```{r parameters}
input_prefix = '/Users/yanyuluchicago/Desktop/tmp/ukb_idp/brainxcan_pipeline/psychiatric_permz/SCZ_PGC_2020'
# meta_file = params$meta_file
color_code_yaml = '/Users/yanyuluchicago/Desktop/tmp/ukb_idp/brainxcan_data/bxcan_vis/report_color_code.yaml'
phenotype_name = 'SCZ_PGC_2020'
# ntop = params$ntop
bxcan_pval_col = 'pval_adj_perm_null'
pval_col_to_name = list(
  pval = 'raw p-values from BrainXcan',
  pval_adj_emp_null = 'adjusted p-values based on simulated weights',
  pval_adj_permz_null = 'adjusted p-values based on permuted weights'
)
outdir = 'new_plots_for_revision'
dir.create(outdir)
```

```{r load_meta}
# gannot = read.csv(meta_file)
color_map = load_color_code_yaml(color_code_yaml)
```

```{r load_results}
bxcan_file = paste0(input_prefix, '.sbrainxcan.csv')
sbxcan = read.csv(bxcan_file) %>% mutate(zscore = abs( qnorm(!!sym(bxcan_pval_col)/2) ) * sign(bhat))
allxcan = sbxcan
```

```{r subset_to_idp_main_results}
allxcan_annot = allxcan # %>% left_join(gannot %>% select(-t1_or_dmri), by = "IDP") 
  
## PC's should have region==PC
ind = substr(allxcan_annot$IDP, 1, 2) == "PC"
allxcan_annot$region[ind] = allxcan$IDP[ind]
## set subtype of PCs as subtype
ind = substr(allxcan_annot$IDP, 1, 2) == "PC"
allxcan_annot$subtype[ind] = allxcan$IDP[ind]

## limit to IDP main results: T1 and TBSS-based dMRI
xcandf = allxcan_annot %>% filter(subtype %in% names(color_map)) %>% rename(side = left_or_right)
```


# BrainXcan results 

In this analysis, we tested `r nrow(allxcan)` structural and diffusion MRI-derived phenotypes with cross validated spearman correlation >= `r signif(min(allxcan$CV_Spearman), digits = 3)`.

```{r, bonforroni_cutoffs}
## define number of total tests per phenotype to be used for bonferroni correction
n_pred_idps = length(unique(xcandf$IDP))
p_thres = 0.05 / n_pred_idps
z_thres = abs(qnorm(p_thres / 2))
```

For the main analysis we focused on `r n_pred_idps` IDPs, which include `r sum(xcandf$subtype == 'Gray-Cortical')` cortical volumes, `r sum(xcandf$subtype == 'Subcortical')` sub-cortical volumes, `r sum(xcandf$subtype == 'Gray-Subcortical')` sub-cortical gray matter volumes.
In below, we show the overall number of IDP by IDP subtype in the main results.

```{r, show_count}
xcandf %>% count(subtype) %>% pander::pander()
```

The p-value being used is `r message(pval_col_to_name[[bxcan_pval_col]])`.

Among `r n_pred_idps` IDPs (`r sum(xcandf$modality=="T1")` T1 and `r sum(xcandf$modality=="dMRI")` diffusion MRI, including the subtype-specific PCs), `r sum(xcandf[[bxcan_pval_col]] < p_thres)` were significantly associated with `r phenotype_name` (pval < $`r signif(p_thres, 3)`$ = 0.05 / `r n_pred_idps`). 

```{r, vis_t1, fig.height=7, fig.width=8}
idp_type = 'T1'
pp = plot_bxcan_ordered(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
```


```{r, vis_dmri, fig.height=6, fig.width=8}
idp_type = 'dMRI'
pp = plot_bxcan_ordered(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
```

# Try heatmap

```{r}
plot_bxcan_ordered_heatmap <- function(xcandf0, color_map, z_thres) {
  shape_map = gen_shape_map()
  tmp = xcandf0 %>% mutate(lr = as.character(side))
  tmp$lr[is.na(tmp$lr)] = 'NA'
  lr2 <- paste0('-', tmp$lr)
  lr2[lr2 == '-NA'] <- ''
  region2 <- tmp$region
  region2[substr(region2, 1, 2) == 'PC'] <- 'PC'
  region2 <- paste0(region2, lr2)
  subtype2 <- tmp$subtype
  pcs <- subtype2[substr(subtype2, 1, 2) == 'PC'] 
  pcs <- unlist(lapply(strsplit(pcs, '-'), function(x) x[2]))
  subtype2[substr(subtype2, 1, 2) == 'PC'] <- pcs
  tmp = tmp %>%
    mutate(region2 = region2, subtype2 = subtype2) %>%
    mutate(kk = reorder(region2, zscore ^ 2, FUN = max))
  tmp2 <- tmp %>% filter(abs(zscore) > z_thres)
  pp = tmp %>%  
    ggplot() + 
    geom_tile(aes(x = kk, y = subtype2, fill = zscore)) +
    geom_text(data = tmp2, aes(x = kk, y = subtype2, label = "*"), color = 'white', size = 4) + 
    coord_flip() +
    ylab('BrainXcan z-score') + 
    theme(axis.title.y = element_blank()) + 
    theme(legend.title = element_blank()) +
    scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0, na.value = 'gray')
  pp
}
```

```{r, vis_t12, fig.height=7, fig.width=6}
idp_type = 'T1'
pp = plot_bxcan_ordered_heatmap(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp = pp + ggtitle(paste0(phenotype_name, ' - ', idp_type)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
ggsave(glue::glue('{outdir}/{phenotype_name}.{idp_type}.png'), width = 6, height = 14)
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
```
```{r, vis_dmri2, fig.height=12, fig.width=4}
idp_type = 'dMRI'
pp = plot_bxcan_ordered_heatmap(xcandf %>% filter(modality == idp_type), color_map, z_thres)
pp = pp + ggtitle(paste0(phenotype_name, ' - ', idp_type))
ggsave(glue::glue('{outdir}/{phenotype_name}.{idp_type}.png'), width = 6, height = 12)
# ggsave(paste0(out_prefix, '.brainxcan_zscore.', idp_type, '.png'), pp, height = 6, width = 8)
```

# Increase color scale

```{r}
srcpath = '/Users/yanyuluchicago/Documents/repo/GitHub/brainxcan/brainxcan/vis'
source(paste0(srcpath, '/', 'vis_helper.R'))

library(dplyr)
library(ggplot2)
library(patchwork)
library(oro.nifti)
theme_set(theme_classic(base_size = 12))
options(stringsAsFactors = F)

opt <- list(
  datadir = '/Users/yanyuluchicago/Desktop/tmp/ukb_idp/brainxcan_data/bxcan_vis',
  pval_col = 'pval_adj_perm_null',
  brainxcan = '/Users/yanyuluchicago/Desktop/tmp/ukb_idp/brainxcan_pipeline/psychiatric_permz/SCZ_PGC_2020.sbrainxcan.csv',
  output_prefix = glue::glue('{outdir}/updated.{phenotype_name}')
)

meta_list = readRDS(paste0(opt$datadir, '/meta_plot.rds'))
tags = names(meta_list)

message(paste0('Loading BrainXcan results. pval_col = ', opt$pval_col))
df = read.csv(opt$brainxcan) %>% mutate(zscore = p2z(!!sym(opt$pval_col), bhat))

message('Plotting for all tags.')
p = list(T1 = list(), dMRI = list()) 
h = list(T1 = c(), dMRI = c())
for(tag in tags) {
  message(paste0('Tag = ', tag))
  idp_type = meta_list[[tag]]$type
  ptmp = vis_by_tag(opt$datadir, tag, df, 'zscore')
  ptmp = add_title(ptmp, meta_list[[tag]]$full_name, size = 20, hjust = 0.2)
  p[[idp_type]][[length(p[[idp_type]]) + 1]] = ptmp
  h[[idp_type]] = c(h[[idp_type]], ceiling(length(unique(ptmp$data$direction)) / 3))
}
for(kk in names(p)) {
  npp = length(p[[kk]])
  if(npp > 0) {
    pp = wrap_plots(p[[kk]], ncol = 1, nrow = npp, heights = h[[kk]])
    ggsave(paste0(opt$output_prefix, '.', kk, '.pdf'), pp, height = 3 * sum(h[[kk]]), width = 10)
  }
}
```